---
title: "Pipeline Exercise"
author: "Diana McSpadden"
date: "8/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question I want to answer: 
    of first gifts that were GIVING CARD gifts in the years 2016, 2017, and 2018 
    how do additional gifts from people who initially gave via giving card compare?
    
## Background:

**A Constituent** is a person, organization, fund, foundation, etc that has a relationship with the museum/non-profit.
        A Constituent may have already giving a gift, or they may be a prospective donor.
    
**A Giving Card** is a piece of paper with donation instructions, including an envelope to mail a donation
        at a later date. Giving Cards are provided on site, typically at points of interest at the museum, 
        by tour guides, or by key historic site staff. Donors who give a gift by GIVING CARD are typically 
        very interested guests who enjoyed their experience at the historic site.
        
**Why ask this question?**
        Giving Card practices may have changed between the year, or guests may be different between the years.
        Are giving cards still an effective means to find new donors?
        
    This question will not answer WHY there are differences between the years of repeat giving by first-time
        Giving Card donors, but since this question has never been asked before it can provide a baseline
        and if we know that Giving Card procedures have changed it may tell us whether they became more or 
        less effective.
        
**Data:**
        The data set was extracted to contain constituents that gave a FIRST GIFT via giving card. 
        The data set contains:
            Constituent ID (unique ID for the constituent)
                ** To anonymize: I removed Constituent Name and Org Name
            Org or Person Indicator: because I removed the Person Name, and Org Name columns I added a P, or O
                to tell me if the Constituent is a person or organization
            First Gift Date: date of first gift to historic site
            First Gift Amount: amount of first gift
            2016 Total Giving: total amount GIFT CARD giving in 2016
            2017 Total Giving: total amount GIFT CARD giving in 2016
            2018 Total Giving: total amount GIFT CARD giving in 2016
            2019 Total Giving: total amount GIFT CARD giving in 2016
            2020 Total Giving: total amount GIFT CARD giving in 2016
            Total Giving All Years: Res and Unres: Total giving (restricted and unrestricted)
                from the Constituent in all years of giving
            
**We only have Giving Card data for 2016 - 2020**
            I can really only compare repeat giving in years 2017, 2018, and 2019
            This is not enough information for a real analysis, but it is a start

```{r loadlibraries}
# load packages
library(tidyverse)
library(ggplot2)
library(stringi)
```

```{r readcsv}
# Part 1: Read data into data frame
df <- read_csv("Lifetime_Giving_First_Gift_Giving_Card_Anon.csv", col_names = TRUE)
```
## Rename columns
in R, the columns that start with numbers will cause problems with coding. I will rename those columns

```{r renamecolumns}
names(df)[names(df)=="2016GCGiving"] <- "GCGiving2016"

names(df)[names(df)=="2017GCGiving"] <- "GCGiving2017"

names(df)[names(df)=="2018GCGiving"] <- "GCGiving2018"

names(df)[names(df)=="2019GCGiving"] <- "GCGiving2019"

names(df)[names(df)=="2020GCGiving"] <- "GCGiving2020"

# print the new names
names(df)
```

Do a basic summarise of the dataframe, but beware these means include all the people who didn't donate in each year, i.e. there are a bunch of $0 donations included in mean.

```{r}
# do a basic analysis of the data 
df %>%
  summarise(meanTotalGiving = mean(TotalGivingResUnres ), iqrTotal = IQR(TotalGivingResUnres), 
  mean2016 = mean(GCGiving2016),
  mean2017 = mean(GCGiving2017),  
  mean2018 = mean(GCGiving2018),   
  mean2019 = mean(GCGiving2019), 
  mean2020 = mean(GCGiving2020), 
  n = n())
```

Get the year by year means that filter out the zero donations.
## 2016 Donations

```{r}
df %>%
  filter(GCGiving2016 > 0) %>%
  summarise(mean2016 = mean(GCGiving2016),
  n = n())

df %>%
  filter(GCGiving2017 > 0) %>%
  summarise(mean2017 = mean(GCGiving2017),
  n = n())

df %>%
  filter(GCGiving2018 > 0) %>%
  summarise(mean2018 = mean(GCGiving2018),
  n = n())

df %>%
  filter(GCGiving2019 > 0) %>%
  summarise(mean2019 = mean(GCGiving2019),
  n = n())

df %>%
  filter(GCGiving2020 > 0) %>%
  summarise(mean2020 = mean(GCGiving2020),
  n = n())
```
## Add the first gift year column 

Add the FirstGiftYear column to the data frame so I can create the 

```{r}
df <- mutate(df, FirstGiftYear = stri_sub(FirstGiftDate,-4,-1))
```

## # create column AdditionalDonations
Create a new column AdditionalDonations for TotalGiving > FirstGiving column called AdditionalDonations
```{r}
df <- mutate(df, AdditionalDonations = (TotalGivingResUnres - FirstGiftAmount))
```

## Create our histograms
Create a histograms from First Gift Year and More Donations

```{r}
df %>%
  group_by(FirstGiftYear) %>%
  filter(AdditionalDonations > 0) %>%
  summarise(countAdditionalGiving = n())
```
```{r}
ggplot(data = df, aes(x = FirstGiftYear)) +
  geom_histogram(stat = "count") + 
  ggtitle("Number of Gifts per First Gift Year")
```
## Average Amount of Additional Giving By First Year Gift

```{r}
df %>%
  group_by(FirstGiftYear) %>%
  filter(AdditionalDonations > 0) %>%
  summarise(averageAdditionalGiving = mean(AdditionalDonations))
```
```{r}
# Remove $0 Additional Donations
dfNoZeroAdditionalDonations <- df %>%
  filter(AdditionalDonations > 0)

ggplot(data = dfNoZeroAdditionalDonations, aes(x = FirstGiftYear, y = AdditionalDonations)) +
  stat_summary(fun.y = "mean", geom = "bar") +
  ggtitle("Average Additional Giving By First Year of Giving")
```

